\section{Quran Phonetic Script}
\label{sec:qps}
We consider the Quran Phonetic Script to be the most valuable and important contribution of our work. By formalizing the assessment of Holy Quran pronunciation as an ASR problem represented through this script, we provide a comprehensive solution to the task.

Modern Standard Arabic (MSA) orthography cannot adequately represent Tajweed rules for error detection. For example, MSA cannot measure the precise length of Madd rules. Previous research (e.g., \cite{omran2023automatic}) focused on single rules like Qalqalah. Our phonetic script addresses this limitation by capturing all Tajweed pronunciation errors except Ishmam (\arb{إشمام}), which involves a visual mouth movement without audible output.

We based our script on classical Muslim scholarship rather than the International Phonetic Alphabet (IPA) for these reasons:

\begin{enumerate}
	\item \textbf{Historical Precedence}: Muslim scholars from the 6th to 14th centuries rigorously defined Quranic errors centuries before modern phonetics emerged in the West.
	\item \textbf{Scientific Foundation}: Scholars like Al-Khalil ibn Ahmad (6th century AH) systematically described articulations and attributes with remarkable accuracy comparable to modern phonetics \cite{article-khalil}.
	\item \textbf{Pedagogical Relevance}: Learners' errors align with classical definitions according to expert Quran teachers.
\end{enumerate}

Following \cite{sweed2021}, Quran recitation errors fall into three categories: \textbf{Articulation Errors} (incorrect pronunciation of phonemes by making the phoneme exits not from its exact location), \textbf{Attribute Errors} (mistakes in letter characteristics (Sifat al-Huruf)), and \textbf{Tajweed Rule Errors} (incorrect application of rules like Ghunnah, Madd, etc.).

Our script comprehensively addresses all three aspects through two main output levels: \textbf{Phonemes Level}, representing letters, vowels, and Tajweed rules, and \textbf{Sifat Level}, consisting of 10 levels representing articulation attributes for each phoneme. Refer to Tables~\ref{tab:phoneme_set} and \ref{tab:sifat_set} for Phonemes and Sifat levels.

Our script has some important charactirstics:

\begin{itemize}
	\item Normal Madd appears as consecutive madd symbols (e.g., 4-beat Madd: \arb{اااا})
	\item Madd al-Leen represented with multiple waw/yaa symbols

	\item Stressed Ghunnah (e.g., \arb{النون المشددة}) as three consecutive noon symbols (\arb{ننن})
	\item Ikhfa represented as three consecutive noon\_mokhfah (\arb{ںںں}) or meem\_mokhfah (\arb{۾۾۾})

	\item Assimilation represented by doubling (e.g., \arb{مَن يَعْمَلْ} $\rightarrow$ \arb{مَيييَعمَل})

	\item Sakin: No following symbol
	\item Imala: fatha\_momala and alif\_momala
	\item Rawm: dama\_mokhtalasa marker
\end{itemize}

Table~\ref{tab:examples_with_sifat} illustrates the phonetization process for the word (\arb{أَتُحَٰٓجُّوٓنِّى}), showing its conversion into phonetic components and Sifat attributes; refer to the table caption for a detailed row-by-row breakdown.





\begin{table*}[h]
	\centering
	\caption{Examples of Uthmani to Phonetic Script Conversion with Sifat Attributes. This example shows the phonetization of word (\arb{أَتُحَٰٓجُّوٓنِّى}) row by row: The first row shows conversion of (\arb{أَ}) to (\arb{ءَ}) with its sifat in the row, following the second row. For the fourth row, showing the madd Lazim rule with 6 beats phoneticized as 6 alifs (\arb{اااااا}), same as the sixth row but with damma represented as 6 waw\_madd (\arb{ۥۥۥۥۥۥ}). And for the fifth row we notice that we converted (\arb{جُّ}) to (\arb{ججُ}) disassembling shadda. The last row shows the normal madd of yaa with two beats represented as two yadd\_madd (\arb{ۦۦ})}
	\label{tab:examples_with_sifat}
	\scriptsize
	\setlength{\tabcolsep}{3pt}
	\begin{tabular}{@{}>{\centering\arraybackslash}m{1.2cm} >{\centering\arraybackslash}m{1.5cm} *{10}{>{\centering\arraybackslash}m{0.7cm}@{}}}
		\toprule
		\textbf{Uthmani} & \textbf{Phonetic} &
		\textbf{H/J}     &
		\textbf{S/R}     &
		\textbf{T/T}     &
		\textbf{Itb}     &
		\textbf{Saf}     &
		\textbf{Qal}     &
		\textbf{Tik}     &
		\textbf{Taf}     &
		\textbf{Ist}     &
		\textbf{Gho}                                                                                     \\
		\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-12}
		\arb{أَ}          & \arb{ءَ}           & jahr & shd & mrq & mnf & no & nql & nkr & ntf & nst & nmg \\
		\arb{تُ}          & \arb{تُ}           & hams & shd & mrq & mnf & no & nql & nkr & ntf & nst & nmg \\
		\arb{حَـ}         & \arb{حَ}           & hams & rkh & mrq & mnf & no & nql & nkr & ntf & nst & nmg \\
		\arb{ـٰٓ}          & \arb{اااااا}      & hams & rkh & mrq & mnf & no & nql & nkr & ntf & nst & nmg \\
		\arb{جُّ}          & \arb{ججُ}          & jahr & shd & mrq & mnf & no & nql & nkr & ntf & nst & nmg \\
		\arb{وٓ}          & \arb{ۥۥۥۥۥۥ}      & jahr & rkh & mrq & mnf & no & nql & nkr & ntf & nst & nmg \\
		\arb{نِّ}          & \arb{ننننِ}        & jahr & btw & mrq & mnf & no & nql & nkr & ntf & nst & mg  \\
		\arb{ى}          & \arb{ۦۦ}          & jahr & rkh & mrq & mnf & no & nql & nkr & ntf & nst & nmg \\
		\bottomrule
	\end{tabular}

	\vspace{2mm}
	\begin{center}  % Centering wrapper added here
		\scriptsize
		\textbf{Attribute Abbreviations:} \\
		H/J: Hams/Jahr \quad S/R: Shidda/Rakhawa \quad T/T: Tafkheem/Taqeeq \quad Itb: Itbaq \\
		Saf: Safeer \quad Qal: Qalqla \quad Tik: Tikraar \quad Taf: Tafashie \quad Ist: Istitala \quad Gho: Ghonna \\

		\textbf{Value Abbreviations:} \\
		shd: shadeed \quad rkh: rikhw \quad btw: between \quad mrq: moraqaq \\
		mof: mofakham \quad mnf: monfateh \quad mtb: motbaq \quad no: no\_safeer \\
		nql: not\_moqalqal \quad nkr: not\_mokarar \quad ntf: not\_motafashie \\
		nst: not\_mostateel \quad nmg: not\_maghnoon \quad mg: maghnoon
	\end{center}  % End of centering wrapper
\end{table*}


\subsection{Development Methodology}
Our phonetization has two steps:
\begin{enumerate}
	\item \textbf{Imlaey to Uthmani Conversion} \\
	      We selected Uthmani script as our foundation because:
	      \begin{itemize}
		      \item Contains specialized Tajweed diacritics (Madd, Tasheel, etc.)
		      \item Preserves pause rules critical for recitation (e.g., stopping on \arb{رحمت})
	      \end{itemize}

	      In order to do that, we created an annotation UI to manually annotate misaligned words in both scripts. For example \ref{tab:imlaey_to_uthmani_ex}, after that, we developed an algorithm that relies on the annotations to convert Imlaey to Uthmani.
	\item \textbf{Uthmani to Phonetic Script Conversion} \\
	      We implemented the process through 26 sequential operations. Each operation contains one or more regular expressions, as shown in the Appendix~\ref{subsec:conversion_ops}.

	\item \textbf{Extracting Sifat}: Next, we extract the 10 attributes (Sifat) defined in Table~\ref{tab:sifat_set}, excluding \textbf{Inhiraf} (\arb{إنحراف}), as it describes the shidda/rakhawa spectrum, and \textbf{Leen} (\arb{اللين}), as it was already handled through our Madd representation.
\end{enumerate}

\begin{table}[h]
	\caption{Example of misalignment between Uthmani and Imlaey Scripts}
	\label{tab:imlaey_to_uthmani_ex}
	\centering
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Imlaey Script} & \textbf{Uthmani Script} \\
		\hline
		\arb{يَا ابْنَ أُمَّ}        & \arb{يَبْنَؤُمَّ}             \\
		\hline
	\end{tabular}
\end{table}


