\section{Related Work}

\subsection{Quran Pronunciation Datasets}
Table~\ref{tab:datasets-comparison} summarizes the most relevant Quran pronunciation datasets. EveryAyah is the largest openly available dataset with 26 complete \emph{Moshafs} segmented and annotated Ayah-by-Ayah by both experts such as Al Hossary and non-experts such as Fares Abbad. Qdat \cite{osman2021qdat} contains 1,509 utterances of single specific Ayahs labeled for three rules: Madd, Ghunna, and Ikhfaa. Although the scale is relatively small, it was widely adopted by the community \cite{10092350,shaiakhmetov2025evaluation} due to being open-source. The Tarteel v1 dataset \cite{khan2021tarteel} consists of 25K utterances with diacritics and no Tajweed rules. The latter is the Tarteel private dataset, a massive 9K-hour collection annotated with diacritics without Tajweed rules. The most recent benchmark is IqraaEval \cite{kheir2025towards}, which presents a test set of 2.2 hours from 18 speakers, but uses Modern Standard Arabic (MSA) without Tajweed rules.

\begin{table}[t]
	\centering
	\scriptsize
	\begin{tabular}{p{2.8cm} p{2.3cm} p{2.2cm} p{3.5cm} c}
		\toprule
		\textbf{Dataset}                                                                                                                       & \textbf{Task}            & \textbf{Size}                           & \textbf{Tajweed Rules}                 & \textbf{Sifat} \\
		\midrule
		EveryAyah                                                                                                                              & ASR / Recitation         & 114 reciters                            & None                                   & No             \\
		Qdat \cite{osman2021qdat}                                                                                                              & Tajweed Detection        & 1,509 utterances                        & 3 (Madd, Ghunna, Ikhfaa)               & No             \\
		Tarteel v1 \cite{khan2021tarteel}                                                                                                      & Diacritization           & 25K utterances (68 h)                   & None                                   & No             \\
		Tarteel (private)                                                                                                                      & Diacritization           & 9K hours                                & None                                   & No             \\
		IqraaEval \cite{kheir2025towards}                                                                                                      & Pronunciation Assessment & 2.2 hours                               & None                                   & No             \\
		\texttt{muaalem-annotated-v3} (ours)\footnote{\href{https://huggingface.co/datasets/obadx/muaalem-annotated-v3}{muaalem-annotated-v3}} & ASR + Tajweed Detection  & 848 h, 286K utterances, 22 reciters     & All Hafs Tajweed rules (except Ishmam) & Yes            \\
		{qdat\_bench} (ours)\footnote{\url{https://huggingface.co/datasets/obadx/qdat_bench}}                                                  & ASR + Tajweed Detection        & 159 samples (1 verse, Al-Ma'idah 5:109) & Madd, Ghunna, Ikhfaa, Qalqalah         & Yes            \\
		\bottomrule
	\end{tabular}
	\caption{Comparison of Quran pronunciation datasets. The Sifat column indicates whether the dataset explicitly encodes \emph{Sifat} (attributes of articulation of Holy Quran phonemes) at the phoneme level.}
	\label{tab:datasets-comparison}
\end{table}

None of the datasets above provides explicit, rule-level Tajweed annotations across the full range of Hafs recitation rules. Qdat, the most widely used benchmark, annotates only three rules and limits evaluation to single-verse utterances. This scarcity of open, richly annotated data has been a fundamental bottleneck for building comprehensive pronunciation assessment systems.

\subsection{Quran Pronunciation Models}
To our knowledge, the first work addressing automated pronunciation assessment for the Holy Quran is RDI \cite{sherif2007enhancing}, which built a complete system for detecting pronunciation errors. The work does not specify which errors were included or excluded but mentions testing Qalqala, Idgham, and Iqlab rules. It also omits details on Quranic word phoneticization. Subsequent work continued with \cite{abdou2014computer} and \cite{al2018computer}, using Deep Neural Networks (DNNs) to replace HMMs and improve the system. Many studies rely on modeling phoneme duration for duration-dependent rules like Madd and Ghunna \cite{mohammed2017recognition,alqadasi2023improving}, but use limited datasets and focus on specific verses rather than the entire Quran. Others concentrate on detecting specific rules like Qalqala \cite{10092350} or Ghunna and Madd \cite{shaiakhmetov2025evaluation,10485145}. The work most aligned with our vision of detecting all error types (including Tajweed and \emph{Sifat}/articulation attributes) is \cite{putra2012developing}. Although it relies on HMMs and minimal data, it introduces a multi-level detection system: \emph{Makhraj} (phoneme level) and Tajweed rules level.

A critical pattern across prior work is that studies are broadly limited in three dimensions: (a) \textbf{data scale} --- most systems train on verse-restricted corpora ranging from dozens to a few hundred samples, which limits generalization to unseen Quranic text; (b) \textbf{rule coverage} --- the majority address only one to three rules (e.g., Qalqala, Madd, Ghunna), leaving the majority of Hafs Tajweed rules unexplored; and (c) \textbf{open-source availability} --- with the exception of Qdat, nearly all prior systems release neither code nor annotated data, making reproducibility and comparison impossible. By contrast, this work targets \emph{all} Hafs Tajweed rules (except Ishmam) using a large expert-recitation corpus and releases both code and data.

At this point, Tarteel emerges; though lacking Tajweed rules, they built a robust ASR system for diacritized character detection. They developed a crowd-sourced dataset \cite{khan2021tarteel} of 25K utterances (68 hours), later extended via application users to 9K hours of private annotated data.

\subsection{Pretrained Speech Encoders with Self-Supervised Learning (SSL)}
Speech pretraining began early \cite{hinton2006reducing} but was constrained by the sequential nature of Recurrent Neural Networks (RNNs) \cite{hopfield1982neural}. The rise of Transformers \cite{vaswani2017attention} facilitated greater GPU parallelization, enabling large-scale pretraining. BERT \cite{devlin2019bert} using Masked Language Modeling (MLM) introduced large unsupervised pretraining which has better results on downstream tasks. This soon extended to speech with wav2vec \cite{schneider2019wav2vec} and wav2vec2.0, which added product quantization \cite{baevski2020wav2vec}. Conformer later replaced vanilla Transformers for speech by integrating convolution \cite{gulati2020conformer}. Google's Wav2Vec2-BERT \cite{chung2021w2v} then applied MLM to speech. Finally, Facebook extended Wav2Vec2-BERT pretraining \cite{barrault2023seamless} to 4.5M hours (including 110K Arabic hours), ideal for low-resource language fine-tuning.

While these general-purpose SSL encoders provide strong representations for a wide range of languages and accents, they are not designed with Tajweed-specific phonetic distinctions in mind. Fine-tuning them directly for Quranic pronunciation assessment therefore requires a task-specific output representation that encodes Arabic phoneme inventory and Tajweed rule boundaries. This work bridges that gap by pairing a large SSL encoder with QPS, a phonetic script that makes Tajweed structure explicit at the label level.
